The complex network of trophic interactions in a subAntarctic oceanic Marine Protected Area

### 2024-04-12 MS_ToSubmit/RevisedVersion_3.Rmd
Final version

Many of these MPAs focus on the presence of particularly vulnerable, keystone, or charismatic species, large numbers (or proportions) of endemic species, and/or high biodiversity across taxonomic levels [@Hogg2016]. Indeed, the MPA N-BB was created to protect a potentially sensitive and biodiverse benthic habitat that was only barely known [@Schejter2016; @Falabella2017]. The benthic community is featured by high biomass of vulnerable and fragile species (mainly Porifera, Bryozoa and Cnidaria) that, considered with their environment, meet the characteristics of vulnerable marine ecosystems [@Schejter2021], here defined as sites that present densities of Indicator Taxa of > 10 kg per 1200 m^2^ [@CommissionfortheConservationofAntarcticMarineLivingResourcesCCAMLR2009]. Also, the benthic realm provides habitat to several small-sized species [@Lopez-Gappa2018; @MartinSirito2019; @Schejter2019], and has an important role in the life history of fishes as a food source, refuge and nursery area [@GarciaAlonso2018; @Vazquez2018; @Delpiani2020; @Troccoli2020; @CovattiAle2022; @Fischer2022; @Matusevich2022; @Florencia2023]. The maintenance of this singular community is related to local and regional oceanographic processes, including the circulation of the rich Malvinas current in the area [@Piola1989; @Guerrero1999] and the upwelling and mixing phenomena [@Matano2019]. The input of nutrients from the Malvinas current also supports a diverse plankton community [@Guinder2020]. Given the cooling trend of this current in a warming regional context (i.e. adjacent waters of the Patagonian shelf and Southwest Atlantic Ocean) [@Franco2022], the MPA N-BB might play an essential role as a refuge maintaining not only the biodiversity but also the fisheries' populations of the region [@Franco2020].

@@

The species-level analysis aims to describe the species' role in the food web. For this, we considered the following properties: betweenness $Btw$, closeness $Cl$, trophic similarity $TS$, topological role $TR$, and trophic level $TL$ (Table 1). Topological roles refer to the fact that food webs tend to naturally organise in non-random, modular patterns, where modules are defined as a group of species that interact more frequently among themselves than with species that are not members of the module [@Guimera2005]. This means that groups of prey and predators interact more strongly with each other, than with species belonging to other groups. Consequently, modularity measures how strongly these modules interact with each other compared to their interaction with species from other modules. Species can play different roles in this respect, according to the pattern of interactions within their own module and/or across modules. We computed the topological role for each species, classified as “module hub”, species with a relatively high number of interactions, but most within its own module; “module specialist”, species with relatively few interactions and most within its own module; “module connector”, species with relatively few interactions mainly between modules; and “network connector”, species with high connectivity between and within modules [@Guimera2005]. Refer to the Supplementary Material for the equations used to calculate the species-level properties.

@@

We are indebted to all those experts of the working group 'Banco Burdwood' who humbly provided their knowledge to enhance the quality of the present research. Although most of them are authors of the present work, it is worth to mention the following researchers: Brenda L. Doti (IBBEA, CONICET-UBA; Universidad de Buenos Aires, Argentina), Sofía L. Callá (Museo Argentino de Ciencias Naturales “Bernardino Rivadavia”, Argentina), Sandra Gordillo (IDACOR-CONICET; Universidad Nacional de Córdoba, Argentina), Mariano I. Martinez (Museo Argentino de Ciencias Naturales “Bernardino Rivadavia”, Argentina) and Luciano Padovani (Instituto Nacional de Investigación y Desarrollo Pesquero, INIDEP, Argentina). We thank the MPA Namuncurá - Burdwood Bank administration. Research cruises were funded by national funds by the Law 26.875. This study was funded by Consejo Nacional de Investigaciones Científicas y Técnicas (CONICET) and Agencia Nacional de Promoción Científica y Tecnológica (PICT 2020 SERIEA 01617), Argentina. This work is contribution no. XX of the MPA Namuncurá (Law 26.875).

>>>>

Many of these MPAs focus on the presence of particularly vulnerable, keystone, or charismatic species, large numbers (or proportions) of endemic species, and/or high biodiversity across taxonomic levels [@Hogg2016]. Indeed, the MPA N-BB was created to protect a potentially sensitive and biodiverse benthic habitat that was only barely known [@Schejter2016; @Falabella2017]. The benthic community is featured by high biomass of vulnerable and fragile species (mainly Porifera, Bryozoa and Cnidaria) that, considered with their environment, meet the characteristics of vulnerable marine ecosystems [@Schejter2021], here defined as sites that present densities of Indicator Taxa of > 10 kg per 1200 m^2^ [@CommissionfortheConservationofAntarcticMarineLivingResourcesCCAMLR2009]. Also, the benthic realm provides habitat to several small-sized species [@Lopez-Gappa2018; @MartinSirito2019; @Schejter2019], and has an important role in the life history of fishes as a food source, refuge and nursery area [@GarciaAlonso2018; @Vazquez2018; @Delpiani2020; @Troccoli2020; @CovattiAle2022; @Fischer2022; @Matusevich2022; @Florencia2023]. The persistence of this particular community is linked to the interplay of local and regional oceanographic dynamics, encompassing the influence of the nutrient-rich Malvinas current [@Piola1989; @Guerrero1999], and occurrences of upwelling and mixing. Moreover, the influx of nutrients carried by the Malvinas current fosters a varied planktonic ecosystem [@Guinder2020]. Given the cooling trend of this current in a warming regional context (i.e. adjacent waters of the Patagonian shelf and Southwest Atlantic Ocean) [@Franco2022], the MPA N-BB might play an essential role as a refuge maintaining not only the biodiversity but also the fisheries' populations of the region [@Franco2020].

@@

The species-level analysis aims to describe the species' role in the food web. For this, we considered the following properties: betweenness $Btw$, closeness $Cl$, trophic similarity $TS$, topological role $TR$, and trophic level $TL$ (Table 1). Topological roles refer to the fact that food webs tend to naturally organise in non-random, modular patterns, where modules are defined as a group of species that interact more frequently among themselves than with species that are not members of the module [@Guimera2005]. This implies that prey and predator groups exhibit greater interaction within their respective modules compared to interactions with species outside those modules. Therefore, modularity assesses the intensity of interactions among these modules relative to interactions with species from different modules. Species can play different roles in this respect, according to the pattern of interactions within their own module and/or across modules. We computed the topological role for each species, classified as “module hub”, species with a relatively high number of interactions, but most within its own module; “module specialist”, species with relatively few interactions and most within its own module; “module connector”, species with relatively few interactions mainly between modules; and “network connector”, species with high connectivity between and within modules [@Guimera2005]. Refer to the Supplementary Material for the equations used to calculate the species-level properties.

@@

We are indebted to all those experts of the working group 'Banco Burdwood' who humbly provided their knowledge to enhance the quality of the present research. Although most of them are authors of the present work, it is worth to mention the following researchers: Brenda L. Doti (IBBEA, CONICET-UBA; Universidad de Buenos Aires, Argentina), Sofía L. Callá (Museo Argentino de Ciencias Naturales “Bernardino Rivadavia”, Argentina), Sandra Gordillo (IDACOR-CONICET; Universidad Nacional de Córdoba, Argentina), Mariano I. Martinez (Museo Argentino de Ciencias Naturales “Bernardino Rivadavia”, Argentina) and Luciano Padovani (Instituto Nacional de Investigación y Desarrollo Pesquero, INIDEP, Argentina). Part of the methodology for the network analysis performed in this work is available in the pre-print @Marina2018 on https://www.biorxiv.org/content/10.1101/094557v1.full. We thank the MPA Namuncurá - Burdwood Bank administration. Research cruises were funded by national funds by the Law 26.875. This study was funded by Consejo Nacional de Investigaciones Científicas y Técnicas (CONICET) and Agencia Nacional de Promoción Científica y Tecnológica (PICT 2020 SERIEA 01617), Argentina. This work is contribution no. XX of the MPA Namuncurá (Law 26.875).
